---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Previus configs
  * Done
```{r setup II, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE)
```


## Installing libraries
```{r echo=FALSE, results=FALSE, message=FALSE}
defaultW <- getOption("warn") 
options(warn = -1) 

#install.packages("ggplot2") ## install
library("ggplot2")            ## load
#install.packages("dplyr")   ## install
library("dplyr")            ## load
```
  * Done

## Loading and preprocessing the data
```{r}
unzip("activity.zip")
df <- read.csv2(file="activity.csv",na.strings = "NA",skip = 0,sep = ",")

```



## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
date_steps <-
  df %>%
  group_by(date) %>%
  summarize(steps_by_day = sum(steps))
date_steps$date <- as.POSIXct(date_steps$date)
summary(date_steps)
```

2. Make a histogram of the total number of steps taken each day

```{r echo=TRUE}
g<-ggplot(data=date_steps,aes(x=steps_by_day) )
g+geom_histogram(binwidth = 2000,color = "purple", fill="deepskyblue4")+labs(x = "Steps by Day",y="Days", title = "Histogram",subtitle = "Total number of steps taken each day")

```

3. Calculate and report the mean and median of the total number of steps taken per day
 + mean
```{r }
mean(date_steps$steps_by_day,na.rm = TRUE)
```
 + median
```{r }
median(date_steps$steps_by_day,na.rm = TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?



```{r}
#Return warn original values
options(warn = defaultW)
```

